IGRINS spectra of \name

### 2016-02-22 document/ms.tex
Revise text draft to include verbatim text from proposal.  Todo is to edit this down.

\documentclass[iop,revtex4]{emulateapj}% change onecolumn to iop for fancy, iop to twocolumn for manuscript
%\documentclass[onecolumn]{emulateapj}% change onecolumn to iop for fancy, iop to onecolumn for manuscript
%\documentclass[preprint]{aastex}

@@

\let\pwiflocal=\iffalse \let\pwifjournal=\iffalse
%From: http://arxiv.org/format/1512.00483
\input{setup}

\providecommand{\eprint}[1]{\href{http://arxiv.org/abs/#1}{#1}}
\providecommand{\adsurl}[1]{\href{#1}{ADS}}
\newcommand{\name}{LkCa4}
\def\vsini{$v\sin{i_*}$}

@@

\name is a weak lined \emph{T-Tauri} star.  In this short research note we provide the IGRINS spectrum of \name.

@@

\section{Analysis}
We compared the IGRINS spectrum of \name to the emph{Phoenix} grid of pre-computed synthetic model spectra\citep{}.  We adapted the \emph{Starfish} code \citep{} in the following way.

@@

We ran the code in several different modes.  First we ran the code as described in \citet{}, with XX orders in the $H$ band, covering XX-XX wavelengths.  The stellar parameters derived in this mode are shared among the orders.  In the second mode, we ran Starfish individually on each of XX orders.  In this mode the stellar properties are derived separately for each order.  Table XX lists the values of Teff derived from the two approaches.

@@

We want to measure the derived effective temperature as a function of wavelength.  The motivation for this strategy is as follows.  We assume that LkCa4 has a significant fraction of its stellar disk covered by starspots \cite{XX}.  Such a photosphere will have an emergent spectrum composed of superpositions of relatively cool patches and warm patches.  The visible portion of the spectrum will be dominated by the warm patches, but the contrast with the starspots will be higher.  The infrared spectrum will have a lower overall spot contast, but with higher relative contribution attributable to the cool patches \citep[e.g. Prato]{XX}.

@@

In short, here is what we know about \name.  It is a weak-lined \emph{T-Tauri} star.  It exhibits a 3.36-3.37 day period \citep{1993AJ....106.1608V,1994IBVS.4042....1G}, as evinced by photometric monitoring in $BVRI$ bands.  No periodic signal is detected at $U-$band.  Section 5 of \citet{1993AJ....106.1608V}, ``More on the nature of CTT and WTT spots'' goes into detail about the whether the photometric variability is attributable to hot or cold patches.  \name has no detected M-type-or-earlier binary companion down to $0.13''$ \citep{1993A&A...278..129L}.  \name was demonstrates variability in X-rays with ROSAT \citep{1994ApJ...424..237S}, with $L_{x}=2.4$ ergs/s.

>>>>

%\documentclass[iop,revtex4]{emulateapj}% change onecolumn to iop for fancy, iop to twocolumn for manuscript
\documentclass[onecolumn]{emulateapj}% change onecolumn to iop for fancy, iop to onecolumn for manuscript
%\documentclass[12pt,preprint]{aastex}

@@

\let\pwiflocal=\iffalse \let\pwifjournal=\iffalse
%From: http://arxiv.org/format/1512.00483
%\input{setup}
\input{mgs_setup}

\providecommand{\eprint}[1]{\href{http://arxiv.org/abs/#1}{#1}}
\providecommand{\adsurl}[1]{\href{#1}{ADS}}
\newcommand{\name}{LkCa4 }
%\def\vsini{$v\sin{i_*}$}

@@

Stars form by direct gravitational collapse of molecular cloud cores.  By $\sim$1 Myr the envelopes of gas and dust have usually accreted onto the star, classical \emph{T-Tauri} stars).  By $\sim$3-10 Myr the circumstellar disks are usually gone (Class III, weak-lined \emph{T-Tauri} stars).  At some time in this process, gas giant planets form.  Observations of young stars between 0 and 10 Myr anchor the overall ``cartoon picture'' understanding of star and planet formation.

The understanding of the process is incomplete.  Measurements for intrinsic stellar properties---luminosity, effective temperature, surface gravity---are known only coarsely.  Their uncertainties flow down to uncertainties in the fundamental properties mass and age, which are estimated from positions on pre main-sequence HR diagrams.

For a single presumably coeval stellar cluster, the pre main-sequence HR diagrams exhibit large ($\delta t_{\ast} \sim t_{\ast}$) apparent spreads in age \citep[e.g.][]{2011A&A...534A..83R}. The large spreads in these pre main-sequence HR diagrams is controversial: some take it as evidence for intrinsic age spreads, some take it as evidence for non static accretion history \citep{2009ApJ...702L..27B, 2010ARA&A..48..581S}.

Still others take the large apparent age spreads in pre main-sequence HR diagrams as evidence for the limited physics included in the pre main-sequence evolutionary model isochrones.  Most modern evolutionary models do not yet include the effects of starspots or magnetic fields, despite observations betraying their presence, \emph{e.g.} optical monitoring and Zeeman doppler imaging \citep{2008A&A...479..827G,2014MNRAS.444.3220D}.

Sunspots are probably responsible for biases in stellar effective temperatures derived by different methods.  For example, the APOGEE spectrograph ($1.5-1.70 \;\mu$m at $R=22,500$) measured effective temperatures for 3493 young stars finding offsets in $\teff$ of typically 200$-$500 K and as high as 1000 K compared to previous studies \citep{2014ApJ...794..125C}.  Low resolution optical spectroscopy shows a typical spread of 200 K in the spectral type-to-effective-temperature conversion scale \citep{2014ApJ...786...97H}.  Stellar evolution models including the effect of starspots can make a coeval 10 Myr population exhibit apparent age spreads of 3$-$10 Myr, with derived masses biased towards lower masses.  The observations of sunspots are lagging behind the theory \citep{2015ApJ...807..174S}.

The importance of starspots needs to be empirically evaluated to assess systematic biases of pre-main sequence stellar evolutionary models.

In this paper we characterize the starspot coverage and temperature contrast of \name using full spectrum fitting.  We demonstrate the constraining power of full spectrum fitting with panchromatic optical and near-IR echelle spectroscopy.  We show that relatively small starspot filling factors can be detected due to the large spectral grasp of IGRINS.

@@

\section{Analysis}
We compared the IGRINS spectrum of \name to the emph{Phoenix} grid of pre-computed synthetic model spectra citeXX.  We adapted the \emph{Starfish} code citeXX in the following way.

@@

We ran the code in several different modes.  First we ran the code as described in citeXX, with XX orders in the $H$ band, covering XX-XX wavelengths.  The stellar parameters derived in this mode are shared among the orders.  In the second mode, we ran Starfish individually on each of XX orders.  In this mode the stellar properties are derived separately for each order.  Table XX lists the values of Teff derived from the two approaches.

On November 15, 2015 we acquired IGRINS miniqueue observations of LkCa4, a weak-lined \emph{T-Tauri} star.  We targeted this source for its likelihood of large areal coverage of starspots, as evinced by large-amplitude photometric variability  detected in $BVRI$ bands.  Its period is 3.37$\pm$0.01 days \citep{1993AJ....106.1608V,1994IBVS.4042....1G}.  Recent optical spectro-polarimetry showed evidence for hot or cool starspots covering an estimated $25\%$ of the stellar surface \citep{2014MNRAS.444.3220D}.  The source has little or no accretion ($\log{\dot M} < 8.1$), no detection of mid-infrared nor mm excess, and no stellar companion, so its spectrum should be devoid of complicating factors like near-IR excess veiling and accretion excess.  This source is the ideal candidate for direct measurement of two-temperature photosphere due to starspots.

We combined our IGRINS spectroscopy with existing ESPaDOnS panchromatic optical echelle spectra to provide 88 spectral orders between 5100 and 25000 \AA.  Our analysis procedure involves fitting the entire spectrum with pre-computed stellar model grids \citep{2013A&A...553A...6H}, and employing a new flexible likelihood function for spectroscopic inference \citep{2015ApJ...812..128C}.  The existing implementation \emph{simultaneously} fits both the intrinsic stellar parameters, $\vt_{\ast} = \{\teff, \logg, \Z \}$, and extrinsic stellar parameters, $\vt_{\rm ext} = \{\sigma_v, \vsini, v_r, \Omega, A_V\}$, with all $N_{\mathrm{ord}}$ spectral orders sharing the same stellar parameters\footnote{The implementation \emph{also} simultaneously fits a small-amplitude polynomial to each spectral order to adjust for slight imperferfections in the continuum shape left over from the blaze correction in the IGRINS pipeline.  The coefficients of the polynomials are treated as nuisance parameters.  This strategy makes us resilient against the pitfalls of pseudo-continuum-fitting, and effortlessly propagates the uncertainty attributable to the continuum level into the uncertainty in the astrophysically interesting stellar parameters.}.  We modified the default implementation by chunking the spectrum into 58 of its cleanest spectral orders and deriving the stellar parameters $\vt_{\ast,m}$ \emph{independently} in each $m^{\mathrm{th}}$ order.  We therefore have 58 unique estimates for $\teff$ as a function of wavelength, which we compile in Figure \ref{fig:teffOrder}.  The figure shows a dramatic effect: \textbf{The derived effective temperature is on-average about 800 K cooler in the $K-$band than in the optical.}

We interpret the cooler temperatures derived at $K-$band as the presence of sunspots contributing more flux at longer wavelengths, as shown in the bottom panel of Figure \ref{fig:teffOrder}.  Here we show the absolute flux ratio, $f_{\lambda, B} / f_{\lambda, A}$ at two different temperatures: $T_\textrm{eff,A} = 4100$ and $T_\textrm{eff,B} = 3300$.  We present the ratio of two synthetic spectra from Phoenix (blue solid line), and the black body ratio (red dashed line).

We further performed a plausibility check to constrain the effect size of starspots.  We generated flux calibrated spectra at two temperatures, and forward-modeled them to resemble the LkCa4 IGRINS and ESPaDOnS spectra in all other ways (\emph{i.e.} $\logg, \Z, \vt_{\rm ext}$), including noise and calibration parameters.  We coadded the spectra in a mixture model:  $ \mathsf{M}_{mix} = c \cdot \mathsf{M}_A(T_\textrm{eff,A}) + (1-c) \cdot \mathsf{M}_B(T_\textrm{eff,B})$.  We chose a fill factor of starspots of 30\%.  We then re-ran our single-temperature fitting procedure on the two-temperature, synthetic, noised-up data to see what stellar parameters one would na\"{\i}vely derive.  The noised-up optical data yields $\teff \sim 4100$ K, close to that of the A component.  We are awaiting the results from $K-$ band at the time of writing.

Our ongoing work is focused on several improvements to the whole-spectrum fitting procedure.  First we are adding support for directly fitting a mixture model, $\mathsf{M}_{mix}$.  This will provide a \emph{measurement} of the star spot temperature and areal coverage fraction.  Mixture model fitting capability will enable true whole-spectrum fitting, a more powerful on-label use of the fitting machinery.

@@

We want to measure the derived effective temperature as a function of wavelength.  The motivation for this strategy is as follows.  We assume that LkCa4 has a significant fraction of its stellar disk covered by starspots citeXX.  Such a photosphere will have an emergent spectrum composed of superpositions of relatively cool patches and warm patches.  The visible portion of the spectrum will be dominated by the warm patches, but the contrast with the starspots will be higher.  The infrared spectrum will have a lower overall spot contast, but with higher relative contribution attributable to the cool patches citeXX.

@@

In short, here is what we know about \name.  It is a weak-lined \emph{T-Tauri} star.  It exhibits a 3.36-3.37 day period \citep{1993AJ....106.1608V,1994IBVS.4042....1G}, as evinced by photometric monitoring in $BVRI$ bands.  No periodic signal is detected at $U-$band.  Section 5 of \citet{1993AJ....106.1608V}, ``More on the nature of CTT and WTT spots'' goes into detail about the whether the photometric variability is attributable to hot or cold patches.  \name has no detected M-type-or-earlier binary companion down to $0.13''$ \citep{1993A&A...278..129L}.  \name demonstrates variability in X-rays with ROSAT \citep{1994ApJ...424..237S}, with $L_{x}=2.4$ ergs/s.
